First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-06-27 ms.tex
NT comments

\affilauthorlist{
A.~P.~Beardsley\affils{ASU,email},
B.~J.~Hazelton\affils{UW,UWeSci},
I.~S.~Sullivan\affils{UW},
P.~Carroll\affils{UW},
N.~Barry\affils{UW},
M.~Rahimi\affils{UMelbourne},
B.~Pindor\affils{UMelbourne,CAASTRO},
C.~M.~Trott\affils{Curtin,CAASTRO},
J.~Line\affils{UMelbourne,CAASTRO},
Daniel~C.~Jacobs\affils{ASU},
M.~F.~Morales\affils{UW}, 
J.~C.~Pober\affils{Brown,UW},
G.~Bernardi\affils{SKASA,RU,CfA},
Judd~D.~Bowman\affils{ASU},
M.~P.~Busch\affils{ASU},
F.~Briggs\affils{ANU,CAASTRO},
R.~J.~Cappallo\affils{Haystack}, 
B.~E.~Corey\affils{Haystack}, 
A.~de~Oliveira-Costa\affils{MIT},
Joshua~S.~Dillon\affils{Berkeley, RAL, BCCP,MIT},
D.~Emrich\affils{Curtin},
A.~Ewall-Wice\affils{MIT},
L.~Feng\affils{MIT},
B.~M.~Gaensler\affils{USydney,CAASTRO,Dunlap}, 
R.~Goeke\affils{MIT},
L.~J.~Greenhill\affils{CfA},
J.~N.~Hewitt\affils{MIT},
N.~Hurley-Walker\affils{Curtin},
M.~Johnston-Hollitt\affils{Victoria},
D.~L.~Kaplan\affils{UWisc}, 
J.~C.~Kasper\affils{UMichigan,CfA}, 
HS Kim\affils{UMelbourne,CAASTRO},
E.~Kratzenberg\affils{Haystack}, 
E.~Lenc\affils{USydney,CAASTRO},
A.~Loeb\affils{CfA},
C.~J.~Lonsdale\affils{Haystack}, 
M.~J.~Lynch\affils{Curtin}, 
B.~McKinley\affils{UMelbourne,CAASTRO},
S.~R.~McWhirter\affils{Haystack},
D.~A.~Mitchell\affils{CASS,CAASTRO}, 
E.~Morgan\affils{MIT}, 
A.~R.~Neben\affils{MIT},
N.~Thyagarajan\affils{ASU},
D.~Oberoi\affils{Tata}, 
A.~R.~Offringa\affils{ASTRON,CAASTRO}, 
S.~M.~Ord\affils{Curtin,CAASTRO},
S. Paul\affils{RRI},
T.~Prabu\affils{RRI}, 
P.~Procopio\affils{UMelbourne,CAASTRO},
J.~Riding\affils{UMelbourne,CAASTRO},
A.~E.~E.~Rogers\affils{Haystack}, 
A.~Roshi\affils{NRAO}, 
N.~Udaya~Shankar\affils{RRI}, 
Shiv~K.~Sethi\affils{RRI},
K.~S.~Srivani\affils{RRI}, 
R.~Subrahmanyan\affils{RRI,CAASTRO}, 
M.~Tegmark\affils{MIT},
S.~J.~Tingay\affils{Curtin,CAASTRO}, 
M.~Waterson\affils{Curtin,ANU},
R.~B.~Wayth\affils{Curtin,CAASTRO}, 
R.~L.~Webster\affils{UMelbourne,CAASTRO}, 
A.~R.~Whitney\affils{Haystack}, 
A.~Williams\affils{Curtin}, 
C.~L.~Williams\affils{MIT},
C.~Wu\affils{UWA},
J.~S.~B.~Wyithe\affils{UMelbourne,CAASTRO}
}

@@

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}, and 
\citealt{Pober:2015} was able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Vedantham:2012, Morales:2012, Datta:2010}. Despite this 
setback, the EoR window is preserved above the wedge, and the field is pushing forward with 
several analysis pipelines under active development \citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2013, Trott:2014}.

@@

The \eppsilon package performs a discrete Fourier transform (DFT) on each $r_{||}$ slice of 
the integrated HEALPix cube, forming a cube in $(\mathbf{k}_{\perp},r_{||})$, where $
\mathbf{k}_{\perp}$ represents the cosmological wavenumber in the plane perpendicular to 
our line of sight, and $r_{||}$ is the distance to the observed redshift slice along the line of 
sight. The Fourier transform along the $r_{||}$ dimension is treated separately due to 
incomplete $(u,v)$ sampling and flagged frequency channels, which leads to structure in the 
frequency sampling along any given $k_{\perp}$ pixel. We thus adopt the Lomb-Scargle 
least-squares method to determine the orthogonal eigenfunctions given our sampling 
function and estimate the total power in each $k_{||}$ mode \citep{Scargle:1982}. In 
addition, because the spectrally smooth foregrounds contain vastly more power than the 
expected EoR signal, we apply a Blackman-Harris window function prior to the $r_{||}$ to 
$k_{||}$ transform. This has the effect of trading off lower effective bandwidth for higher 
dynamic range.

@@

\section{Discussion}\label{sec:discussion}
Inspecting our 32 hour integrated power spectra we see two places where we are limited. 
First there is a region that is leakage-dominated at low $k$. The cause of this leakage is yet 
unknown, but likely due to imperfect calibration, beam models, and/or foreground models. 
In order to improve on our limit in this regime, future analysis will require calibration which 
better accounts for the instrumental response, perhaps by increasing signal to noise 
through multiple-snapshot solutions or refined antenna response models. Studies of the 
beam are underway \citep[e.g.][]{Neben:2015}, and will likely help to further isolate the 
foreground contaminates in our power spectra. The foreground model is a constant area of 
investigation, but improved spectral dependence in both the point source catalog and the 
diffuse model, as well as the addition of an all-sky Galactic model, will continue to improve 
the foreground subtraction and further unlock the EoR window.

>>>>

\affilauthorlist{
A.~P.~Beardsley\affils{ASU,email},
B.~J.~Hazelton\affils{UW,UWeSci},
I.~S.~Sullivan\affils{UW},
P.~Carroll\affils{UW},
N.~Barry\affils{UW},
M.~Rahimi\affils{UMelbourne},
B.~Pindor\affils{UMelbourne,CAASTRO},
C.~M.~Trott\affils{Curtin,CAASTRO},
J.~Line\affils{UMelbourne,CAASTRO},
Daniel~C.~Jacobs\affils{ASU},
M.~F.~Morales\affils{UW}, 
J.~C.~Pober\affils{Brown,UW},
G.~Bernardi\affils{SKASA,RU,CfA},
Judd~D.~Bowman\affils{ASU},
M.~P.~Busch\affils{ASU},
F.~Briggs\affils{ANU,CAASTRO},
R.~J.~Cappallo\affils{Haystack}, 
B.~E.~Corey\affils{Haystack}, 
A.~de~Oliveira-Costa\affils{MIT},
Joshua~S.~Dillon\affils{Berkeley, RAL, BCCP,MIT},
D.~Emrich\affils{Curtin},
A.~Ewall-Wice\affils{MIT},
L.~Feng\affils{MIT},
B.~M.~Gaensler\affils{USydney,CAASTRO,Dunlap}, 
R.~Goeke\affils{MIT},
L.~J.~Greenhill\affils{CfA},
J.~N.~Hewitt\affils{MIT},
N.~Hurley-Walker\affils{Curtin},
M.~Johnston-Hollitt\affils{Victoria},
D.~L.~Kaplan\affils{UWisc}, 
J.~C.~Kasper\affils{UMichigan,CfA}, 
HS Kim\affils{UMelbourne,CAASTRO},
E.~Kratzenberg\affils{Haystack}, 
E.~Lenc\affils{USydney,CAASTRO},
A.~Loeb\affils{CfA},
C.~J.~Lonsdale\affils{Haystack}, 
M.~J.~Lynch\affils{Curtin}, 
B.~McKinley\affils{UMelbourne,CAASTRO},
S.~R.~McWhirter\affils{Haystack},
D.~A.~Mitchell\affils{CASS,CAASTRO}, 
E.~Morgan\affils{MIT}, 
A.~R.~Neben\affils{MIT},
Nithyanandan~Thyagarajan\affils{ASU},
D.~Oberoi\affils{Tata}, 
A.~R.~Offringa\affils{ASTRON,CAASTRO}, 
S.~M.~Ord\affils{Curtin,CAASTRO},
S. Paul\affils{RRI},
T.~Prabu\affils{RRI}, 
P.~Procopio\affils{UMelbourne,CAASTRO},
J.~Riding\affils{UMelbourne,CAASTRO},
A.~E.~E.~Rogers\affils{Haystack}, 
A.~Roshi\affils{NRAO}, 
N.~Udaya~Shankar\affils{RRI}, 
Shiv~K.~Sethi\affils{RRI},
K.~S.~Srivani\affils{RRI}, 
R.~Subrahmanyan\affils{RRI,CAASTRO}, 
M.~Tegmark\affils{MIT},
S.~J.~Tingay\affils{Curtin,CAASTRO}, 
M.~Waterson\affils{Curtin,ANU},
R.~B.~Wayth\affils{Curtin,CAASTRO}, 
R.~L.~Webster\affils{UMelbourne,CAASTRO}, 
A.~R.~Whitney\affils{Haystack}, 
A.~Williams\affils{Curtin}, 
C.~L.~Williams\affils{MIT},
C.~Wu\affils{UWA},
J.~S.~B.~Wyithe\affils{UMelbourne,CAASTRO}
}

@@

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}, and 
\citealt{Pober:2015} was able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Thyagarajan:2013, Vedantham:2012, Morales:2012, Datta:2010}. Despite this 
setback, the EoR window is preserved above the wedge, and the field is pushing forward with 
several analysis pipelines under active development \citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2013, Trott:2014}.

@@

The \eppsilon package performs a discrete Fourier transform (DFT) on each $r_{||}$ slice of 
the integrated HEALPix cube, forming a cube in $(\mathbf{k}_{\perp},r_{||})$, where $
\mathbf{k}_{\perp}$ represents the cosmological wavenumber in the plane perpendicular to 
our line of sight, and $r_{||}$ is the distance to the observed redshift slice along the line of 
sight. The Fourier transform along the $r_{||}$ dimension is treated separately due to 
incomplete $(u,v)$ sampling and flagged frequency channels, which leads to structure in the 
frequency sampling along any given $k_{\perp}$ pixel. We thus adopt the Lomb-Scargle 
least-squares method to determine the orthogonal eigenfunctions given our sampling 
function and estimate the total power in each $k_{||}$ mode \citep{Scargle:1982}. In 
addition, because the spectrally smooth foregrounds contain vastly more power than the 
expected EoR signal, we apply a Blackman-Harris window function prior to the $r_{||}$ to 
$k_{||}$ transform. This has the effect of trading off lower effective bandwidth for higher 
dynamic range \citep[e.g.][]{Thyagarajan:2016, Thyagarajan:2013}.

@@

\section{Discussion}\label{sec:discussion}
Inspecting our 32 hour integrated power spectra we see two places where we are limited. 
First there is a region that is leakage-dominated at low $k$. The cause of this leakage is yet 
unknown, but likely due to imperfect calibration, beam models, and/or foreground models.
In order to improve on our limit in this regime, future analysis will require calibration which 
better accounts for the instrumental response, perhaps by increasing signal to noise 
through multiple-snapshot solutions or refined antenna response models. 
Through simulation, \citealt{Thyagarajan:2016} recently demonstrated the necessity of
 precise foregrounds and instrumental models in order to retain a clean EoR window.
Studies of the 
beam are underway \citep[e.g.][]{Neben:2015}, and will likely help to further isolate the 
foreground contaminates in our power spectra. The foreground model is a constant area of 
investigation, but improved spectral dependence in both the point source catalog and the 
diffuse model, as well as the addition of an all-sky Galactic model, will continue to improve 
the foreground subtraction and further unlock the EoR window.
